---
title: "Practical Record: Business Analytics for Agriculture"
author: "Name: Vijay,  Roll No: 1234"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-location: body
    toc-title: "Table of Contents"
    number-sections: true
    toc-depth: 3
    code-copy: true
editor_options: 
  chunk_output_type: console
---

---

```{r}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
knitr::opts_chunk$set(message = FALSE)
```

---

# Introduction to R Programming

## Basic R Scripts

Below are a few examples of basic R scripts that demonstrate common tasks in R.

### Example 1: Print "Hello World"

A simple script that prints "Hello, World!" to the console.

```{r}
# Print "Hello, World!" to the console
print("Hello, World!")
```

## Basic Operators 

### Assignment Operator (`=` or `<-`)
Use `<-` or `=` for assigning values to variables

```{r}
x=5
x
y <- 3
y
```

### Commenting Code for Clarity

Use `#` for comments. Comments are not executable and are used to provide relevant information about the syntax

```{r}
# Assigning values to x
x=5
x
```

### Arithmetic operators

- Primary arithmetic operators available in R  are Addition `+` , Subtraction `-`, Multiplication `*`, Division `/`, Sqaure `^`

```{r}
# Perform arithmetic operations
add <- 5 + 3

# Print the results
add
```

**Square (\^) operator** - Squares the first number by the second.

```{r}
x=5
x^2
```

## Statements

### Logical Operations

**Equality**: `==` checks if two values are equal.

```{r}
# Equality 5 == 3
x <- 5
y <- 3
x == y  
```

**Inequality**: `!=` checks if two values are not equal.

```{r}
# Inequality 5 != 3
x != y  
```

**Greater than**: `>` checks if the value on the left is greater than the value on the right.

```{r}
# Greater than 5 > 3
x > y   
```

**Less than**: `<` checks if the value on the left is less than the value on the right.

```{r}
# Less than 5 < 3
x < y   
```

**Greater than or equal to**: `>=` checks if the value on the left is greater than or equal to the value on the right.

```{r}
# Greater than or equal to 5 >= 3
x >= y  
```

**Less than or equal to**: `<=` checks if the value on the left is less than or equal to the value on the right.

```{r}
# Less than or equal to 5 <= 3
y <= x 
```

## Data Structures

### Vectors

-   Vectors are one-dimensional arrays that can store numeric, character, or logical data.
-   The `c()` function combines values into a vector. 

```{r, filename="creating vector"}
# Numeric vector
age <- c(20, 21, 23)
age
# Character vector
student <- c("ajay", "vijay", "jay")
student
# logical vector
Pass<- c(TRUE, FALSE, TRUE) 
Pass
```

### Matrix

-   Created using the matrix() function. nrow is used to set number of rows and byrow is used to set values by rows (if TRUE) or columns (if FALSE).

```{r}
# Create a matrix with 2 rows and 3 columns
my_matrix <- matrix(data = c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3, byrow = TRUE)

# Print the matrix
print(my_matrix)
```

### Array

-   Similar to matrices but can have more than two dimensions.
-   Created using the array() function. dimensions are set using dim.

```{r}
# Create a 3-dimensional array with dimensions 3x3x2
my_array <- array(data = c(1:18), dim = c(3, 3, 2))

# Print the array
print(my_array)
```

---

## Functions

Consists inbuilt functions like `sum()`, `length()`, `sqrt()`,`mean()`, `summary()`, `View()`

### Arithmetic Functions

#### sum() Function

The `sum()` function calculates the total sum of all the elements in a numeric vector.

```{r, filename="sum function"}
# Calculate the sum of two variables
x=5
y=3
z=2
sum(x,y,z)
# Calculate the sum of a numeric vector
numbers <- c(1, 2, 3)
sum(numbers)
```

#### length() Function

The `length()` function returns the number of elements in a vector (or other objects).

```{r, filename="length function"}
# Find the length of a vector
numbers <- c(1, 2, 3, 4, 5)
length(numbers)
```

#### sqrt() Function

The `sqrt()` function calculates the square root of each element in a numeric vector.

```{r, filename="square root function"}
# Calculate square root of a variable
x=25
sqrt(x)
# Calculate the square root of each element in a numeric vector
numbers <- c(1, 4, 9, 16, 25)
sqrt(numbers)
```

### mean() Function

The `mean()` function calculates the arithmetic mean (average) of the elements in a numeric vector.

```{r,  filename="mean function"}
# Example: Calculate the mean of a numeric vector
numbers <- c(2,4,6)
mean(numbers)
```

### summary() function

The `summary()` function in R provides a concise statistical summary of objects like vectors, matrices, data frames, and results of model fitting.

```{r,  filename="summary function"}
# Create a numeric vector
vec <- c(1, 2, 3, 4, 5, NA, 7, 8, 9, 10)

# Get summary of the vector
summary(vec)
```

### data.frame() function

`data.frame()` function is used to create data frames, which are table-like structures consisting of rows and columns. 

```{r, filename="creating data frame"}
# Creating a simple data frame
students <- data.frame(
  Name = c("Arun", "Arun", "Charan", "Divya", "Eswar", 
           "Fathima", "Gopal", "Harini", "Ilango", "Jayanthi",
           "Kiran", "Lavanya", "Mohan", "Nandini", "Omkar", 
           "Pavithra", "Qasim", "Raji", "Sanjay", "Tulsi"),
  Age = c(25, 25, 35, 28, 22, 
          40, 33, 27, 31, 29,
          24, 26, NA, 28, 30,
          27, 34, 29, NA, 26),
  Height = c(5.6, 5.6, NA, 5.4, 6.0, 
             5.3, 5.9, 5.5, 5.7, 5.8,
             5.9, NA, 6.1, 5.4, 5.6,
             5.5, 5.8, 5.3, 5.7, NA),
  Gender = c("M", "M", NA, "F", "M", 
             "F", "M", "F", "M", "F",
             "M", "F", "M", NA, "M",
             "F", "M", "F", "M", "F"),
  Marks = c(75, 75, 79, 80, 81, 
            82, 83, 84, 85, 86,
            87, 88, 83, 82, 85, 
            81, 79, 86, 84, 78),
  Attendance = c(92, 92, 88, 90, 87, 
                 95, 83, 89, 86, 91,
                 90, 88, 85, 89, 93,
                 87, 86, 91, 84, 90),
  Residence = c("Day Scholar", "Day Scholar", "Hosteller", "Hosteller", "Day Scholar",
                "Hosteller", "Hosteller", "Day Scholar", "Hosteller", "Day Scholar",
                "Hosteller", "Day Scholar", "Hosteller", "Day Scholar", "Hosteller",
                "Hosteller", "Day Scholar", "Hosteller", "Day Scholar", "Day Scholar")
)

students
```

### head() function

The head() function in R is used to display the first few rows of a dataset.

```{r, filename="creating data frame"}
head(students)
```

### View() function

`View() function` is used to invoke a spreadsheet-like data viewer on a data frame, matrix, or other objects.

```{r, filename="View() function", eval=FALSE}
# View function to see the data in a dedicated window
View(students)
```

---

### Loops

Use `for`, `while`.

#### for loop

The for loop in R is used to iterate over a sequence (like a vector or a list) and execute a block of code for each element in the sequence.

```{r, filename="for loop"}
# Print numbers from 1 to 5
for(i in 1:5) {
  print(i)
}
```

#### while loop

The while loop executes a block of code as long as the specified condition is TRUE

```{r, filename="while loop"}
# Print numbers from 1 to 5 using while
i <- 1  # Initialize counter
while(i <= 4) {
  print(i)
  i <- i + 1  # Increment counte
}
```

---

## Directories and Projects in R

### Setting a New Working Directory: 
The `setwd()` function allows users to change the working directory.

```{r, eval=FALSE}
setwd("/Users/vijay/Library/")
```

### Current Working Directory:
After setting a new directory, you can verify it by running `getwd()`.

This helps users confirm where R is looking for files and where outputs will be saved.

```{r}
# Get the current working directory
getwd()
```

---

## Overview of Key R Packages

- `stats`: Statistical analysis and hypothesis testing  
- `plotly`: Interactive visualizations  
- `tidyverse`: A collection of packages for data manipulation and visualization 

### stats: Statistical Analysis and Hypothesis Testing in R

The `stats` package comes pre-installed with R and provides essential statistical functions for data analysis.

#### Basic Usage:
```{r}
# Descriptive statistics
summary(mtcars$mpg)

# t-test example
t.test(mtcars$mpg ~ mtcars$am)

# Linear regression
model <- lm(mpg ~ wt, data = mtcars)
summary(model)
```

### plotly: Creating Interactive Visualizations in R

The `plotly` package enables the creation of **interactive and dynamic charts**.

```{r, eval=FALSE}
# Install and load the package
install.packages("plotly")
```

```{r}
library(plotly)

# Create a simple scatter plot
plot_ly(data = mtcars, x = ~wt, y = ~mpg, type = "scatter", mode = "markers")
```

### Tidyverse: collection of data science packages

The **tidyverse** is a collection of R packages designed for **data science**. It provides a structured approach to **importing, manipulating, visualizing, and modeling** data.

**Installing and Loading Tidyverse**

```{r, eval=FALSE}
# Install tidyverse
install.packages("tidyverse")

# Load all core tidyverse packages
library(tidyverse)
```

This loads several useful packages such as **ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, and forcats**.

#### Data Manipulation with dplyr

```{r}
library(dplyr)

# Sample dataset
data <- tibble(Name = c("John", "Alice", "Bob"), Score = c(85, 90, 78))

# Filtering rows where Score > 80
filtered_data <- data %>% filter(Score > 80)

# Summarizing average score
average_score <- data %>% summarize(Avg_Score = mean(Score))

print(filtered_data)
print(average_score)
```

#### Data Tidying with tidyr

```{r}
library(tidyr)

# Sample dataset (wide format)
data <- tibble(Name = c("A", "B"), Math = c(85, 90), Science = c(78, 92))

# Convert to long format
long_data <- pivot_longer(data, cols = c(Math, Science), names_to = "Subject", values_to = "Marks")

print(long_data)
```

#### Data Visualization with ggplot2

```{r}
library(ggplot2)

# Sample dataset
data <- tibble(x = c(1, 2, 3, 4, 5), y = c(2, 4, 6, 8, 10))

# Create a scatter plot
ggplot(data, aes(x = x, y = y)) +
  geom_point(color = "blue") +
  labs(title = "Scatter Plot", x = "X Values", y = "Y Values")
```

#### String Manipulation with stringr

```{r}
library(stringr)

text <- "Welcome to R programming!"

# Convert to lowercase
lower_text <- str_to_lower(text)

# Replace "R" with "Tidyverse"
new_text <- str_replace(text, "R", "Tidyverse")

print(lower_text)
print(new_text)
```

#### Handling Factors with forcats

```{r}
# Load the library
library(forcats)
library(ggplot2)

# Sample categorical data: Fruit purchases
fruit <- factor(c("Apple", "Banana", "Apple", "Cherry", "Banana", "Banana", 
                  "Apple", "Cherry", "Cherry", "Banana", "Apple"))

# View original levels and frequency
table(fruit)

```

---

# Data Preparation and Transformation in R

**Data Handling in R** refers to the **importing, managing, transforming, and cleaning** of data before analysis. 


## Importing Data into R

- R supports importing data from various sources, including ** Excel, CSV, databases(SQL), JSON, and APIs**.

### Import csv file from website

- use `read.csv()` function.
- Sample csv [website link](https://raw.githubusercontent.com/yuvijen/Datasets/refs/heads/main/Diabetes.csv)

```{r}
# Read csv from website
Diabetes <- read.csv("https://raw.githubusercontent.com/yuvijen/Datasets/refs/heads/main/Diabetes.csv")

# Display the first few rows
head(Diabetes)
```

### Import csv file from local folder

**Sample csv data file** 

- [Download the csv dataset from this link](https://raw.githubusercontent.com/yuvijen/Datasets/refs/heads/main/Agriculture%20production%20india.csv).

- copy the file path name or file name (if working directory or project is set) of the dataset and paste it inside the `read.csv("file name")`

```{r}
# Read CSV file
Indiaagriculture = read.csv("Agri production india.csv")

# Display the first few rows
head(Indiaagriculture)
```

### Import excel file from local folder

To read Excel files in R, install the `readxl` package.

**Install readxl package**

```{r, eval=FALSE}
install.packages("readxl")
```

**Sample data file**

[Download the excel dataset from this link](https://github.com/yuvijen/Datasets/raw/refs/heads/main/Europeanagriculture.xlsx).

- Save this excel file in your project or working directory. 
- copy the file name (or file path name) of the dataset and paste it inside the function `read_excel("file_name")`

```{r}
library(readxl)
Europeanagriculture = read_excel("Europeanagriculture.xlsx", sheet = 1)
head(Europeanagriculture)
```


## Data file creation in R

use `data.frame()` function
```{r}
# Create a simple data frame
students <- data.frame(
  Name = c("Arun", "Arun", "Charan", "Divya", "Eswar", 
           "Fathima", "Gopal", "Harini", "Ilango", "Jayanthi",
           "Kiran", "Lavanya", "Mohan", "Nandini", "Omkar", 
           "Pavithra", "Qasim", "Raji", "Sanjay", "Tulsi"),
  Age = c(25, 25, 35, 28, 22, 
          40, 33, 27, 31, 29,
          24, 26, NA, 28, 30,
          27, 34, 29, NA, 26),
  Height = c(5.6, 5.6, NA, 5.4, 6.0, 
             5.3, 5.9, 5.5, 5.7, 5.8,
             5.9, NA, 6.1, 5.4, 5.6,
             5.5, 5.8, 5.3, 5.7, NA),
  Gender = c("M", "M", NA, "F", "F", 
             "F", "M", "F", "M", "F",
             "M", "F", "M", NA, "M",
             "F", "M", "F", "F", "M"),
  Marks = c(58, 58, 65, 99, 89, 
            80, 81, 82, 83, 83,
            84, 84, 85, 85, 86, 
            96, 87, 98, 97, 92),
  Attendance = c(84, 84, 85, 85, 86, 
                 86, 86, 87, 89, 89,
                 89, 90, 90, 91, 81,
                 92, 83, 95, 96, 94),
  Residence = c("Day Scholar", "Day Scholar", "Hosteller", "Hosteller", "Day Scholar",
                "Hosteller", "Hosteller", "Day Scholar", "Hosteller", "Day Scholar",
                "Hosteller", "Day Scholar", "Hosteller", "Day Scholar", "Hosteller",
                "Hosteller", "Day Scholar", "Hosteller", "Day Scholar", "Day Scholar")
)

# Gender and Residence are categorical variables, so convert them to factor
students$Gender <- factor(students$Gender, 
                 levels = c("M", "F"), 
                 labels = c("Male", "Female"))
students$Residence <- factor(students$Residence, 
                 levels = c("Hosteller", "Day Scholar"), 
                 labels = c("Hosteller", "Day Scholar"))

students
```

---

## Viewing and Exploring Data


### Viewing data structure

- `View()` - To view the entire table in a window
- `head()` - View the first few rows
- `str()` - Check structure of dataset
- `summary()` - Summary statistics

```{r}

head(students)

str(students)

summary(students)
```

### Checking Missing Values
```{r}
# Count missing values per column
colSums(is.na(students))
```

### Checking Duplicates
```{r}
# Identify duplicate rows
duplicates <- students[duplicated(students), ]
duplicates
```

### Handling missing values and imputations

#### Replace missing values with mean

```{r}

# Create a copy of the original data for data cleaning. 

clean <- students
#  Calculate the mean height (excluding NA)
mean_height <- mean(clean$Height, na.rm = TRUE)

# Replace NA values in Height with the calculated mean
clean$Height[is.na(clean$Height)] <- mean_height

clean
```

#### Remove missing values

```{r}
clean <- na.omit(clean)

# Reset row numbering
rownames(clean) <- NULL

clean
```
#### Removing Duplicates

```{r}
clean <- clean[!duplicated(clean), ]

# Reset row numbering
rownames(clean) <- NULL

clean
```

### Changing Data Types
```{r}
# Convert column to numeric
clean$Height <- as.numeric(clean$Height)

# Convert column to character
clean$Name <- as.character(clean$Name)
str(clean)
```

### Data Manipulation with `dplyr`

#### Selecting Specific Columns

- Install dplyr package

```{r, eval=FALSE}
install.packages("dplyr")
```

```{r}
library(dplyr)

# Select specific columns
clean <- clean %>% select(Name,Age,Residence, Gender, Marks, Attendance)
clean
```

#### Removing Columns
```{r}
# Remove specific columns
clean <- clean %>% select(-Residence)
clean
```

#### Filtering Data
```{r}
# Filter rows where column value is greater than 100
clean <- clean %>% filter(clean$Attendance >= 85)
clean
```

#### Sorting Data
```{r}
# Arrange data in descending order
clean <- clean %>% arrange(desc(clean$Marks))
clean
```

### Reshaping Data with `tidyr`


#### Converting Wide Data to Long Format
```{r}
library(tidyr)

clean <- clean %>% pivot_longer(cols = c("Marks", "Attendance"), names_to = "Measure",values_to = "Value")
clean
```

#### Converting long data to wide format

```{r}
# Assuming your current dataframe is named 'clean' in long format:
clean <- clean %>% pivot_wider(names_from = Measure, values_from = Value)

clean
```


###  Exporting Data from R

#### Export the data file as a csv file to local folder


`write.csv(data, "folder path/filename.csv")`
To save the CSV in a specific folder, provide the path.

```{r}
# Export the data frame as csv
write.csv(clean, "cleanstudentsdata.csv")
```

#### Export data file as a excel file to local folder

The `writexl` package provides an easy way to export data from R into an Excel file.

**Install writexl package**

```{r, eval=FALSE}
install.packages("writexl")
```

- export the clean data file as xlsx file. 

```{r}
library(writexl)
# Write data to an Excel file
write_xlsx(clean, "cleanstudentsdata.xlsx")
```

### Normalization

```{r}

# Sample dataset
df <- data.frame(Height = c(150, 160, 170, 180, 190, 200),
                   Weight = c(50, 60, 65, 75, 80, 90))

# Min-Max Normalization function
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# Apply normalization to dataset
df_normalized <- as.data.frame(lapply(df, normalize))

# Print results
df_normalized
```


### Standardization

```{r}
# Standardization function (Z-score scaling)
standardize <- function(x) {
  return ((x - mean(x)) / sd(x))
}

# Apply standardization
df_standardized <- as.data.frame(lapply(df, standardize))

# Print results
df_standardized
```

### Creating Dummy Variables

#### Using `ifelse()` for Binary Categories

If a variable has two categories (e.g., Male/Female), we can create a dummy variable using `ifelse()`.

```{r}
# Sample dataset
  data <- data.frame(Gender = c("Male", "Female", "Male", "Female", "Male"))

# Create a dummy variable for "Male" (1 = Male, 0 = Female)
  data$Male_Dummy <- ifelse(data$Gender == "Male", 1, 0)

print(data)
```

#### Using `model.matrix()` for Multiple Categories

For categorical variables with more than two categories, one-hot encoding is used to create multiple dummy variables.


```{r}
# Sample dataset
  data <- data.frame(Department = c("HR", "IT", "Finance", "HR", "IT"))

# Create dummy variables
  dummy_vars <- model.matrix(~ Department - 1, data = data)

# Convert to dataframe
  dummy_vars_df <- as.data.frame(dummy_vars)

# Print dataset
  print(dummy_vars_df)
```

---

# Data visualization

## 2D data visualization techniques

### Scatter Plot
In R, you can use the `plot()` function from the base package to create a scatter plot.

```{r}

# Create a scatter plot
plot(clean$Marks, clean$Attendance, main="Scatter Plot of Marks vs. Attendance", 
     xlab="Marks", ylab="Attendance", pch=19, col="red")
```


### Histogram
In R, you can use the `hist()` function from the base package to create a histogram.

```{r}
# Vector of scores
# Create a histogram
hist(clean$Age, breaks=5, col="blue", main="Histogram of Marks", xlab="Marks", ylab="Frequency")
```

### Bar Charts

#### Simple Bar Chart

```{r}

# Calculate average marks by gender
avg_marks <- tapply(clean$Marks, clean$Gender, mean)

# Create barplot
barplot(avg_marks,
        col = "navy",
        main = "Average Marks by Gender",
        ylab = "Average Marks",
        xlab = "Gender")

```

#### Sub-divided (Stacked) Bar Chart

```{r}


# Stacked Bar Chart
barplot(rbind(clean$Attendance, clean$Marks), beside=FALSE, col=c("yellow", "red"), 
        legend.text=c("Attendance", "Marks"), names.arg=clean$Name,
        main="Stacked Bar Chart", ylab="Quantity")
```

#### Multiple Bar Chart

```{r}

# Stacked Bar Chart
barplot(rbind(clean$Attendance, clean$Marks), beside=TRUE, col=c("yellow", "red"), 
        legend.text=c("Attendance", "Marks"), names.arg=clean$Name,
        main="Stacked Bar Chart", ylab="Quantity")
```

### Line Plot

```{r}
library(ggplot2)

# Create a data frame with monthly sales data
data <- data.frame(
  Month = factor(c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun'), levels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun')),
  Sales = c(200, 210, 190, 215, 205, 230)
)

# Plotting the line graph
ggplot(data, aes(x=Month, y=Sales)) +
  geom_line(group=1, colour="blue") +
  geom_point() +
  labs(title="Monthly Sales Data", x="Month", y="Sales") +
  theme_minimal()
```

### Pie Chart

```{r}
# Summarize total marks for each gender by grouping 'Marks' based on 'Gender'
gender_marks <- tapply(clean$Marks, clean$Gender, sum, na.rm = TRUE)

percent <- round(gender_marks / sum(gender_marks) * 100)
labels <- paste(names(gender_marks), percent, "%")

# Plot pie chart with value labels
pie(gender_marks,
    labels = labels,
    col = c("skyblue", "pink"),
    main = "Total Marks by Gender")
```

## 3D data visualization techniques

### 3D Pie Chart

```{r}

library(plotly)

# Data
categories <- c("Apple", "Banana", "Cherry")
values <- c(300, 450, 250)

# 3D Pie Chart using plotly
fig <- plot_ly(labels = categories, values = values, type = 'pie', textinfo = 'label+percent',
               insidetextorientation = 'radial', marker = list(colors = c('#FF9999', '#66B3FF', '#99FF99')))
fig <- fig %>% layout(title = "3D Pie Chart", showlegend = TRUE)
fig
```



### 3D Scatter Plot

```{r}
library(plotly)
students <- students[!is.na(students$Gender), ]

# Convert Gender to numeric (e.g., Male = 1, Female = 2)
students$Gender_Num <- as.numeric(as.factor(students$Gender))

# Create 3D scatter plot
plot_ly(students, x = ~Age, y = ~Marks, z = ~Gender_Num, type = "scatter3d", mode = "markers",
        marker = list(size = 5, color = ~Gender_Num, colorscale = "Viridis")) %>%
  layout(title = "3D Scatter Plot of Students",
         scene = list(
           xaxis = list(title = "Age"),
           yaxis = list(title = "Marks"),
           zaxis = list(title = "Gender (Numeric)")
         ))
```

---

# Descriptive Analytics

## Measures of Central Tendency

**Mean**

```{r}
# Define the vector of exam scores
exam_scores <- c(85, 90, 78, 92, 85)

# Calculate the mean (average) score
mean_score <- mean(exam_scores)

# Print the result
print(mean_score)
```

**Median**

```{r}

# Define the vector of exam scores
exam_scores <- c(85, 90, 78, 92, 85)

# Calculate the median score
median_score <- median(exam_scores)

# Print the result
print(median_score)
```

**Mode**

```{r}

# Define the vector of exam scores
exam_scores <- c(85, 90, 78, 92, 85)

# Compute mode using which.max()
mode_score <- as.numeric(names(which.max(table(exam_scores))))

# Print the result
print(mode_score)
```


## Measures of Dispersion

```{r}
# Define the dataset
data <- c(1, 2, 4, 7, 9)

# Calculate Range
data_range <- max(data) - min(data)

# Calculate Interquartile Range (IQR)
data_iqr <- IQR(data)

# Calculate Variance
data_variance <- var(data)

# Calculate Standard Deviation
data_sd <- sd(data)

# Calculate Mean Deviation (Mean Absolute Deviation)
data_mad <- mean(abs(data - mean(data)))

# Print Results
print(paste("Mean Deviation:", data_mad))
print(paste("Standard Deviation:", data_sd))
print(paste("Variance:", data_variance))
print(paste("Range:", data_range))
print(paste("Interquartile Range (IQR):", data_iqr))
```


## Measures of Skewness and Kurtosis

**Normal distribution skewness**
```{r}
# Load the library
library(moments)

# Define the dataset
exam_scores <- c(50, 51, 52, 54, 55, 56, 57, 58, 59, 60, 60, 61, 62, 63, 63, 64, 65, 66, 67, 67, 68, 69, 70, 72, 73, 74, 75, 78)

# Calculate Skewness
exam_skewness <- skewness(exam_scores)

# Print the result
print(exam_skewness)

```


**Left skewed**

```{r}
# Load the library
library(moments)

# Define the dataset
exam_scores <- c(50, 53, 55, 55, 56, 57, 58, 58, 59, 59, 60, 60, 61, 61, 61, 62, 62, 62, 63, 63, 63, 64, 64, 64, 65, 65, 66, 66)

# Calculate Skewness
exam_skewness <- skewness(exam_scores)

# Print the result
print(exam_skewness)

```

**Right skewed**

```{r}
# Load the library
library(moments)

# Define the dataset
exam_scores <- c(5, 7,8, 9, 10, 11, 12, 13, 14, 15, 16,17, 18, 19, 20, 21, 22, 23, 24,25, 26, 27, 28, 30, 32, 35, 38, 40, 42)

# Calculate Skewness
exam_skewness <- skewness(exam_scores)

# Print the result
print(exam_skewness)

```


**Histogram**

```{r}
# Load necessary libraries
library(ggplot2)
library(moments)


# Calculate skewness
exam_skewness <- skewness(exam_scores)

# Create histogram with density curve
ggplot(data.frame(exam_scores), aes(x = exam_scores)) +
  geom_histogram(aes(y = after_stat(density)), bins = 5, fill = "blue", color = "black", alpha = 0.6) +
  geom_density(color = "red", linewidth = 1) +
  labs(title = paste("Histogram and Density Curve of Exam Scores\nSkewness:", round(exam_skewness, 2)),
       x = "Exam Scores",
       y = "Density") +
  theme_minimal()

```


**Kurtosis**
```{r}
# Load the library
library(moments)

# Define the dataset
exam_scores <- c(55, 60, 65, 65, 70, 75, 80)

# Calculate Skewness
exam_kurt <- kurtosis(exam_scores)

# Print the result
print(exam_kurt)

```

**Histogram**

```{r}
# Load necessary libraries
library(ggplot2)
library(moments)

# Define the dataset
exam_scores <- c(55, 60, 65, 65, 70, 75, 80)

# Create a dataframe
exam_data <- data.frame(Score = exam_scores)

# Plot histogram with density curve
ggplot(exam_data, aes(x = Score)) +
  geom_histogram(aes(y = after_stat(density)), bins = 5, fill = "blue", alpha = 0.5, color = "black") +
  geom_density(color = "red", linewidth = 1) +
  labs(title = "Histogram and Density Curve of Exam Scores",
       x = "Exam Scores",
       y = "Density") +
  theme_minimal()

```

## Measures of Association

### Covariance
```{r}
# Covariance between MPG and Horsepower
cov(mtcars$mpg, mtcars$hp)
```


### Correlation
```{r}
# Correlation between MPG and Horsepower
cor(mtcars$mpg, mtcars$hp)  
```

#### Scatter Plot
```{r}
# Load necessary library
library(ggplot2)

# Scatter plot with regression line
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(color = "blue", size = 3, alpha = 0.7) +  # Scatter plot points
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # Regression line
  labs(title = "Scatter Plot of MPG vs Horsepower",
       x = "Horsepower (hp)",
       y = "Miles Per Gallon (mpg)") +
  theme_minimal()

```

#### Positive Correlation
```{r}
cor(students$Attendance , students$Marks)  # Correlation 

```

```{r}

# Load necessary library
library(ggplot2)

# Scatter plot with regression line
ggplot(students, aes(x = Attendance, y = Marks)) +
  geom_point(color = "blue", size = 3, alpha = 0.7) +  # Scatter plot points
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # Regression line
  labs(title = "Scatter Plot",
       x = "Attendance",
       y = "Marks") +
  theme_minimal()

```


#### Negative Correlation

```{r}
students <- students[!is.na(students$Age), ]
cor(students$Age , students$Attendance)  # Correlation 
```

```{r}
# Load necessary library
library(ggplot2)

# Scatter plot with regression line
ggplot(students, aes(x = Age, y = Attendance)) +
  geom_point(color = "blue", size = 3, alpha = 0.7) +  # Scatter plot points
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # Regression line
  labs(title = "Scatter Plot",
       x = "Age",
       y = "Attendance") +
  theme_minimal()
```

---

# Inferential Statistical Techniques

## T-test

### One-Sample T-Test

```{r, filename = "R"}
# Sample data
weights <- c(52, 55, 61, 54, 58, 59, 62, 53, 56, 57, 60, 59, 61, 64, 58)
alpha = 0.05
# Calculate sample size
sample_size <- length(weights)
sample_size
# Calculate sample mean
sample_mean <- mean(weights)
sample_mean 
# Calculate sample standard deviation (R uses n-1 by default)
sample_sd <- sd(weights)
sample_sd
# Define population mean for comparison
population_mean <- 60
population_mean

# Calculate the t-statistic
t_statistic <- (sample_mean - population_mean) / (sample_sd / sqrt(sample_size))
t_statistic
# Degrees of freedom
degrees_of_freedom <- sample_size - 1
degrees_of_freedom
# Calculate the p-value for a two-tailed test
p_value <- 2 * pt(-abs(t_statistic), df = degrees_of_freedom)
p_value
# Decision based on p-value
if (p_value < alpha) {
  cat("Reject null hypothesis\n")
} else {
  cat("Do not reject null hypothesis\n")
}
```


### Two-Sample T-Test 

```{r, filenname = "R"}
# Test scores for two independent classes
class_a_scores <- c(85, 88, 90, 95, 78)
class_b_scores <- c(80, 83, 79, 92, 87)
alpha = 0.05
# Perform independent samples t-test
# We assume equal variances for this example with var.equal = TRUE
t_test_result <- t.test(class_a_scores, class_b_scores, var.equal = TRUE)

# Print the results
t_test_result
p_value = t_test_result$p.value
if (p_value < alpha) {
  cat("Reject null hypothesis\n")
} else {
  cat("Do not reject null hypothesis\n")
}

```


### Paired Samples T-Test

```{r, filename = "R"}
# Participant weights before and after the diet
weights_before <- c(70, 72, 75, 80, 78)
weights_after <- c(68, 70, 74, 77, 76)
alpha = 0.05
# Perform paired samples t-test
t_test_result <- t.test(weights_before, weights_after, paired = TRUE)

# Print the results
t_test_result
# Extract p-value
p_value = t_test_result$p.value
# hypothesis decision
if (p_value < alpha) {
  cat("Reject null hypothesis\n")
} else {
  cat("Do not reject null hypothesis\n")
}
```

---

## ANOVA (Analysis of Variance)

### One way ANOVA Test

```{r}
# Prepare the Data
emp_referral <- c(11, 15, 18, 19, 22)
job_portals <- c(17, 18, 21, 22, 27)
consultancy <- c(15, 16, 18, 19, 22)
alpha = 0.05
# Combining the data into a single data frame
data <- data.frame(
  Sales = c(emp_referral, job_portals, consultancy),
  Method = factor(rep(c("Emp Referral", "Job Portals", "Consultancy"), each = 5))
)
data
# Perform ANOVA Test
result <- aov(Sales ~ Method, data = data)

# Results
summary(result)
# Get the summary of the ANOVA test
summary_result <- summary(result)

# Extract the p-value
p_value <- summary_result[[1]]["Method", "Pr(>F)"]

# hypothesis decision
if (p_value < alpha) {
  cat("Reject null hypothesis\n")
} else {
  cat("Do not reject null hypothesis\n")
}
```


### Two way ANOVA Test in R

```{r, eval=FALSE}
install.packages("stats")
install.packages("agricolae")
```

```{r}
# Loading necessary library
library(stats)
library(agricolae)

# Preparing the data
PlantHeight <- c(15, 17, 16, 14, 15, 15, 18, 20, 19, 22, 21, 23)
FertilizerType <- factor(rep(c('A', 'B'), each=6))
IrrigationMethod <- factor(rep(c('X', 'Y', 'X', 'Y'), each=3))
alpha = 0.05
data <- data.frame(PlantHeight, FertilizerType, IrrigationMethod)
data
# Conducting Two-Way ANOVA
result <- aov(PlantHeight ~ FertilizerType * IrrigationMethod, data = data)
summary(result)

```

---

## Chi-square

### Chi-square Test of Independence

```{r}
# Data for the Chi-square Test of Independence
data <- matrix(c(20, 10, 30, 40), nrow = 2, byrow = TRUE)

# Perform the test
result <- chisq.test(data)

# Output the result
print(result)
```

### Chi-square Goodness-of-Fit Test

```{r}
# Observed sales
observed <- c(50, 60, 40, 50)

# Expected sales (equal distribution)
expected <- rep(sum(observed) / length(observed), length(observed))

# Perform the test
result <- chisq.test(observed, p = expected / sum(expected))

# Output the result
print(result)
```

# Regression 

**Simple Linear Regression**

```{r}  
# Sample dataset  
data <- data.frame(  
  study_hours = c(2, 3, 5, 7, 9),  
  exam_score  = c(50, 55, 65, 80, 90)  
)  

# Build a linear regression model  
model <- lm(exam_score ~ study_hours, data = data)  

# Summary of the model  
summary(model)  
```


**Visualizing the relationship**

```{r}
plot(data$study_hours, data$exam_score, col = "blue", pch = 19, main = "Study Hours vs Exam Score")  
abline(model, col = "red")  
```

**Linear Regression**
```{r}
# Load necessary library
library(ggplot2)

# Sample dataset
house_data <- data.frame(
  Size = c(1500, 1800, 2100, 2500, 1300, 1700, 2200, 2700, 1600, 1400, 
           1900, 2300, 2800, 2900, 2000, 2400, 3000, 2600, 3100, 3200, 
           3300, 3400, 3500, 3600, 3700, 3800, 3900, 4000, 4100, 4200),
  Price = c(300, 340, 400, 450, 260, 320, 420, 480, 310, 280, 
            360, 430, 490, 510, 370, 440, 520, 460, 530, 550, 
            570, 590, 610, 630, 650, 670, 690, 710, 730, 750)
)

# Fit linear regression model
model <- lm(Price ~ Size, data = house_data)

# Summary of the model
summary(model)
```


**Plot the data and regression line**

```{r}
ggplot(house_data, aes(x = Size, y = Price)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "House Size vs Price Regression",
       x = "House Size (sq.ft)",
       y = "House Price ($1000s)")
```

---

# Lab Exam

## Load the data1 dataset
download the dataset from this [link](https://1drv.ms/x/c/ff525616fe275280/EYBSJ_4WVlIggP8sgwMAAAABDihUL4pwAZ7eeOv2c17_tA?e=sZf4lD)
```{r}
library(readxl)
data1 = read_excel("Organic_Products_Survey_Final.xlsx")
str(data1)
```

## Install dplyr package
```{r, eval=FALSE}
install.packages("dplyr")
```

## Rename the variable names
```{r}
library(dplyr)
data1 = data1 %>% rename('Age' = 'Age Group')
data1 = data1 %>% rename('Gender'='Gender')
data1 = data1 %>% rename('Education'='Education Level')
data1 = data1 %>% rename('Income'='Income Level')

data1 <- data1 %>% rename(
  'HC1' = 'I reflect about my health a lot.'
)

data1 <- data1 %>% rename(
  'HC2' = 'I am very self-conscious about my health.'
)

data1 <- data1 %>% rename(
  'HC3' = 'I am alert to changes in my health.'
)

data1 <- data1 %>% rename(
  'HC4' = 'I am usually aware of my health.'
)

data1 <- data1 %>% rename(
  'HC5' = 'I take responsibility for the state of my health.'
)

data1 <- data1 %>% rename(
  'EC1' = 'I am concerned about the environment.'
)

data1 <- data1 %>% rename(
  'EC2' = 'Humans must live in harmony with nature to survive.'
)

data1 <- data1 %>% rename(
  'EC3' = 'I am willing to reduce my consumption to help protect the environment.'
)

data1 <- data1 %>% rename(
  'EC4' = 'Environmental protection is an important issue to me.'
)

data1 <- data1 %>% rename(
  'EC5' = 'I would be willing to pay more for environmentally friendly products.'
)

data1 <- data1 %>% rename(
  'PI1' = 'I intend to purchase organic farm products in the near future.'
)

data1 <- data1 %>% rename(
  'PI2' = 'I will make an effort to buy organic farm products.'
)

data1 <- data1 %>% rename(
  'PI3' = 'I am likely to purchase organic farm products.'
)

data1 <- data1 %>% rename(
  'PI4' = 'I plan to buy organic farm products regularly.'
)

data1 <- data1 %>% rename(
  'PI5' = 'I will recommend organic farm products to others.'
)

```

## Convert categorical variables into factors

```{r}
data1$Age = factor(data1$Age, levels = c( "18-25 Years",	"26-35 Years", "36-45 Years",	"46-55 Years",	"Above 55 Years"), ordered = TRUE)

data1 = data1 %>% mutate('Gender' = factor(Gender, 
                      levels = c("Male", "Female")))
data1 = data1 %>% mutate('Education' = factor(Education, 
                         levels = c("High School", "Diploma", 
                                    "Undergraduate", "Postgraduate", "Doctorate"), 
                         ordered = TRUE))

data1 = data1 %>% mutate('Income' = factor(Income, 
                      levels = c("0 - 2  Lakhs", "2 - 5  Lakhs", 
                                 "5 - 10  Lakhs", "Above 10 Lakhs"), 
                      ordered = TRUE))

```

## Recode likert sclae to Interval or numeric (1-5)

```{r}
data1 <- data1 %>% mutate_at(
  c('HC1', 'HC2', 'HC3', 'HC4', 'HC5', 
    'EC1', 'EC2', 'EC3', 'EC4', 'EC5',
    'PI1', 'PI2', 'PI3', 'PI4', 'PI5'), 
  list(~ recode(., 
    'Strongly Disagree' = 1, 
    'Disagree' = 2, 
    'Neutral' = 3, 
    'Agree' = 4, 
    'Strongly Agree' = 5))
)
```

## Export the coded data

```{r, eval=FALSE}
install.packages("writexl")
```

```{r}
library(writexl)
write_xlsx(data1, "data2.xlsx")
```

```{r}
data2 = read_excel("data2.xlsx")
str(data2)
```


## Descriptive statistics of demographic variables
```{r}
summary(data1$Age)
summary(data1$Gender)
summary(data1$Education)
summary(data1$Income)
```

## Creating Study variables by rowmeans
```{r}
# Compute the mean for Health Consciousness
data1 <- data1 %>%
  mutate(Health_Consciousness = rowMeans(select(., HC1, HC2, HC3, HC4, HC5), na.rm = TRUE))

# Compute the mean for Environmental Concern
data1 <- data1 %>%
  mutate(Environmental_Concern = rowMeans(select(., EC1, EC2, EC3, EC4, EC5), na.rm = TRUE))

# Compute the mean for Purchase Intention
data1 <- data1 %>%
  mutate(Purchase_Intention = rowMeans(select(., PI1, PI2, PI3, PI4, PI5), na.rm = TRUE))
```

## Descriptives for Study variables

```{r}
mean(data1$Health_Consciousness)
sd(data1$Health_Consciousness)
var(data1$Health_Consciousness)
mean(data1$Environmental_Concern)
sd(data1$Environmental_Concern)
var(data1$Environmental_Concern)
mean(data1$Purchase_Intention)
sd(data1$Purchase_Intention)
var(data1$Purchase_Intention)

summary(data1$Health_Consciousness)
summary(data1$Environmental_Concern)
summary(data1$Purchase_Intention)

```

## T Test

### Conduct t-test for Health Consciousness, Environment concern, and Purchase Intention on Gender.

```{r}
t_test_HC <- t.test(data1$Health_Consciousness ~ data1$Gender, data = data1, var.equal = TRUE)
t_test_HC

# Conduct t-test for Environmental Concern
t_test_EC <- t.test(data1$Environmental_Concern ~ data1$Gender, data = data1, var.equal = TRUE)
t_test_EC

# Conduct t-test for Purchase Intention
t_test_PI <- t.test(data1$Purchase_Intention ~ data1$Gender, data = data1, var.equal = TRUE)
t_test_PI

```

## Anova
### ANOVA for Health Consciousness across Education Levels

```{r}
anova_HC_Edu <- aov(Health_Consciousness ~ Education, data = data1)
summary(anova_HC_Edu)

# ANOVA for Environmental Concern across Education Levels
anova_EC_Edu <- aov(Environmental_Concern ~ Education, data = data1)
summary(anova_EC_Edu)

# ANOVA for Purchase Intention across Education Levels
anova_PI_Edu <- aov(Purchase_Intention ~ Education, data = data1)
summary(anova_PI_Edu)

```

## Correlation

```{r}
cor_matrix <- cor(data1[, c("Health_Consciousness", "Environmental_Concern", "Purchase_Intention")], method = "pearson")
cor_matrix
```

## Regression

```{r}
# Run Multiple Linear Regression
regression_model <- lm(Purchase_Intention ~ Health_Consciousness + Environmental_Concern, data = data1)

summary(regression_model)
```

---

